Parameter estimation

### 2023-03-24 SFE-article/Article_Model_Parameters.tex
16.03
24.03

\subsection{Governing equations}
The detail derivation of the governing equation can be found in the appendix (\ref{CH: Gouverning equations}) as well as in the work of \citet{Anderson1995}. Let's assume that any properties of the flow are unifrom across any given cross-section of an channel (or any device like an extractor). Such a flow is called quasi-one-dimensional. The variation of the cross-section might be an result of its irregular shape (or partial filling of an extractor).

@@

{\footnotesize
\begin{align}
		\label{EQ: CompressibleEuler_1}
        \cfrac{\partial \left( \rho_f A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v \right)}{\partial z} &= 0 \\
        \cfrac{\partial \left( \rho_f v A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v^2 \right)}{\partial z} &= -A_f \cfrac{\partial P}{\partial z} {EQ: CompressibleEuler_2} \\
        \cfrac{\partial \left( \rho_f e A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v e\right)}{\partial z} &= -P\cfrac{\left( A_f v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial T}{\partial z} \right)   
        \label{EQ: CompressibleEuler_3}
    \end{align}  
}

>>>>

\subsection{Governing equations}
The detail derivation of the governing equation can be found in the appendix \ref{CH: Gouverning equations} as well as in the work of \citet{Anderson1995}. Let's assume that any properties of the flow are unifrom across any given cross-section of an channel (or any device like an extractor). Such a flow is called quasi-one-dimensional. The variation of the cross-section might be an result of its irregular shape (or partial filling of an extractor).

@@

{\footnotesize
\begin{align}
		\label{EQ: CompressibleEuler_1}
        \cfrac{\partial \left( \rho_f A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v \right)}{\partial z} &= 0 \\
        \cfrac{\partial \left( \rho_f v A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v^2 \right)}{\partial z} &= -A_f \cfrac{\partial P}{\partial z} \label{EQ: CompressibleEuler_2} \\
        \cfrac{\partial \left( \rho_f e A_f \right) }{\partial t} + \cfrac{\partial \left( \rho_f A_f v e\right)}{\partial z} &= -P\cfrac{\left( A_f v \right)}{\partial z} + \cfrac{\partial}{\partial z} \left( \cfrac{\partial T}{\partial z} \right)   
        \label{EQ: CompressibleEuler_3}
    \end{align}  
}

### 2023-03-24 SFE-article/Sections/Model.tex
24.03

		\subsubsection{Heat balance} \label{CH: heat_balance} \todo{I will be back. The energy equation is not given as a in terms of temperature any more.}

		 {\color{red}
			The heat balance (equation \ref{Model_heat}) consists of the convective and diffusive terms. It follows the assumption of a pseudo-homogeneous phase, which properties are the mean between fluid and solid phases. We consider no heat loss through the wall, and there is no heat generation in the system. Therefore, the temperature of the extractor can be changed only by manipulating the temperature of the inlet stream %$T_{In}(t)$.

@@

		\todo[inline]{Draft of new heat equation starts here. TODO: \\ Check if the heat cpacity of solid phase needs to be added \\ Elaborate why pressure is know}

		\noindent\colorbox{pink}
		{
		\begin{minipage}{\columnwidth}
			The heat equation was introduced in the previous chapter through equation \ref{EQ: CompressibleEuler_3} or

@@

			For real gases it is complicated to write the heat balance in terms of temperature. Alternatively, the energy equation can be used in the form of internal energy, so as it is given by equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature (which is used as the input to some other functions) needs to be recover from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by equation \ref{EQ: internal_energy_PR}

@@

			{\footnotesize
			\begin{equation*}
				e(t,z)=e(T(t,z),P(t),\rho(t,z))=e(T(t,z),P(t),\underbrace{\rho(T(t,z),P(t))}_{E.O.S.}) 
			\end{equation*}
			}

@@

		\end{minipage} }

>>>>

		\subsubsection{Heat balance} \label{CH: heat_balance}

		\iffalse
		 {\color{red}
			The heat balance (equation \ref{Model_heat}) consists of the convective and diffusive terms. It follows the assumption of a pseudo-homogeneous phase, which properties are the mean between fluid and solid phases. We consider no heat loss through the wall, and there is no heat generation in the system. Therefore, the temperature of the extractor can be changed only by manipulating the temperature of the inlet stream %$T_{In}(t)$.

@@

%		\todo[inline]{Draft of new heat equation starts here. TODO: \\ Check if the heat cpacity of solid phase needs to be added \\ Elaborate why pressure is know}
		\fi

			The heat equation was introduced in the previous chapter through equation \ref{EQ: CompressibleEuler_3} or

@@

			For real gases it is complicated to write the heat balance in terms of temperature, but it be used directly in the form of internal energy, as it is given by equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature (which is used as the input to some other functions) needs to be recover from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by equation \ref{EQ: internal_energy_PR}.

@@

			{\footnotesize
			\begin{equation*}
				e(t,z)=e(T(t,z),P(t),\rho(t,z))=e( T(t,z), P(t), \rho(T(t,z),P(t)) ) 
			\end{equation*}
			}

@@

			Another way to express the energy equation is to introduce enthalpy $e = h - P_0 / \rho$. By introducing the definition of enthalpy, the energy equation becomes

			{\footnotesize
				\begin{equation}
					\cfrac{\partial \left(\rho_f h A_f\right)}{\partial t} - \cfrac{\partial \left(P_0 A_f\right)}{\partial t} + \cfrac{\partial \left( \rho_f h A_f v \right)}{\partial z} - \cfrac{\partial}{\partial z} \left( k \cfrac{\partial T}{\partial z} \right)
				\end{equation}
			}

			The main advantage of this formulation is the presence of term $\partial P_0 / \partial t $, which allow to directly affect the system through the change of thermodynamic pressure (which is a control variable).\todo{ "For high pressures, the influence of the intermolecular forces on the enthalpy has to be taken into account, usually by applying a cubic equation of state. In most cases, these forces are attractive, so that additional energy is necessary to move the molecules away from each other, that is, to lower the density. If this energy is not added, the substance cools down when it is expanded" (\citet{Gmehling2019}). So, pressure reduction cause gas expansion and temperature drop.}
			If an equation of state is known, the temperature can to be recovered from the enthalpy. The enthalpy is related to the pressure and temperature through the following equation:

			{\footnotesize
				\begin{equation}
					h(t,z)=h(T(t,z),P(t),\rho(t,z))=h( T(t,z), P(t), \rho(T(t,z),P(t)) ) 
					\label{EQ:Enthalpy_root}
				\end{equation}
			}

			If the value of enthalpy is know from the time evolution and pressure can be measured, then the equation \ref{EQ:Enthalpy_root} can be solved for temperature to recover the temperature profile. By applying the departure functions to Peng-Robinson equation of state, the relation \ref{EQ:Enthalpy_root} can be expressed directly through equation \ref{EQ:Enthalpy_PR} as presented in \ref{CH:Enthalpy} or given by \citet{Gmehling2019}.

			{\footnotesize
				\begin{equation}
					h_{T,P}-h_{T,P}^{\mathrm {ideal} }=RT_{C}\left[T_{r}(Z-1)-2.078(1+\kappa ){\sqrt {\alpha }}\ln \left({\frac {Z+2.414B}{Z-0.414B}}\right)\right]
					\label{EQ:Enthalpy_PR}
				\end{equation}
			}

			The equation \ref{EQ:Enthalpy_PR} requires an reference sate, which in this case is assumed to be $T_{ref}=298.15[K]$ and $P_{ref}=1.01325[bar]$.

### 2023-03-24 SFE-article/Sections/Thermo.tex
24.03

    where $a, b, \alpha$ are parameters defined as presented in the appendix.

    The properties of the $CO_2$ presented as a function of operating conditions (temperature and pressure) are presented on fig. \ref{fig: SFE_Properties}

@@

    where $A$ and $B$ are parameters as defined in the appendix (\ref{CH: Thermodynamic_details}). The roots of the polynomial can be found iteratively or by Cardano formula. Depends on the operating conditions, one or two roots can be found. In one-phase region the fluid can be describe as gas, liquid or super-critical. In the two-phase region the gas-liquid mixture is present. The biggest root is assigned to the gas phase and smallest root corresponds to the liquid phase.

    The real gas effects are also visible on the density plot presented on the Figure \ref{fig: SFE_Properties_Density}. The density can change significantly depends on the operating conditions. 
    By analysing the compressibility and density plots more it can be noticed that very near to the critical point, the fluid can neither be called a liquid nor a gas and has a unique combination of gas-like and liquid-like properties.

    The Figure \ref{fig: SFE_Properties_CP} show behaviour of the heat capacity of a supercritical fluid at constant pressure ($C_p$). The details of the calculations can be found in the appendix. In contrary to the density which varies monotonically, the specific heat shows very high levels in a narrow region. In the subcritical region, the phase transition is associated with an effective spike in the heat capacity (i.e., the latent heat). Approaching the critical point, the latent heat falls to zero but this is accompanied by a gradual rise in heat capacity in the pure phases near phase transition. At the critical point, the latent heat is zero but the heat capacity shows a diverging singularity. Beyond the critical point, there is no divergence, but rather a smooth peak in the heat capacity; the highest point of this peak identifies the Widom line (as discussed by \citet{Simeoni2010} and \citet{Banuti2019}).

>>>>

    where $a, b, \alpha$ are parameters defined as presented in the appendix \ref{subsubsec: Equation of state}.

    The properties of the $CO_2$ presented as a function of operating conditions (temperature and pressure) are presented on Figure \ref{fig: SFE_Properties}

@@

    where $A$ and $B$ are parameters as defined in the appendix \ref{subsubsec: Equation of state}. The roots of the polynomial can be found iteratively or by the Cardano formula. One or three real roots can be found depending on the operating conditions. In a one-phase region, the fluid is described by one real root corresponding to the gas, liquid or supercritical phase. In the two-phase region, the gas-liquid mixture is present and has two roots. The biggest root is assigned to the gas phase, and the smallest root corresponds to the liquid phase.

 	The real gas effects are also visible on the density plot presented on the Figure \ref{fig: SFE_Properties_Density}. The density calculations are based on the compressibility factor, as explained in the appendix \ref{subsubsec: Fluid density}, and its value depends on the operating conditions.

    By analysing the compressibility and density plots more it can be noticed that very near to the critical point, the fluid can neither be called a liquid nor a gas and has a unique combination of gas-like and liquid-like properties.

    The Figure \ref{fig: SFE_Properties_CP} show behaviour of the heat capacity of a supercritical fluid at constant pressure ($C_p$). The details of the calculations can be found in the appendix \ref{subsubsec: Fluid heat capacity}. In contrary to the density which varies monotonically, the specific heat shows very high levels in a narrow region. In the subcritical region, the phase transition is associated with an effective spike in the heat capacity (i.e., the latent heat). Approaching the critical point, the latent heat falls to zero but this is accompanied by a gradual rise in heat capacity in the pure phases near phase transition. At the critical point, the latent heat is zero but the heat capacity shows a diverging singularity. Beyond the critical point, there is no divergence, but rather a smooth peak in the heat capacity; the highest point of this peak identifies the Widom line (as discussed by \citet{Simeoni2010} and \citet{Banuti2019}).

### 2023-03-24 SFE-article/Sections/Thermodynamic_details.tex
24.03

	\todo[]{Add internal energy equation}

>>>>

	\subsubsection{Departure functions for enthalpy calculations} \label{CH:Enthalpy} \todo[]{Change to my text}

	In thermodynamics, a departure function is defined for any thermodynamic property as the difference between the property as computed for an ideal gas and the property of the species as it exists in the real world, for a specified temperature T and pressure P. Common departure functions include those for enthalpy, entropy, and internal energy.

	Departure functions are used to calculate real fluid extensive properties (i.e. properties which are computed as a difference between two states). A departure function gives the difference between the real state, at a finite volume or non-zero pressure and temperature, and the ideal state, usually at zero pressure or infinite volume and temperature.

	For example, to evaluate enthalpy change between two points $h(V_1,T_1)$ and $h(V_2,T_2)$ we first compute the enthalpy departure function between volume $V_1$ and infinite volume at $T = T_1$, then add to that the ideal gas enthalpy change due to the temperature change from $T_1$ to $T_2$, then subtract the departure function value between $V_2$ and infinite volume.

	Departure functions are computed by integrating a function which depends on an equation of state and its derivative. The general form the enthalpy equation is given by

	{\footnotesize 
		\begin{equation}
			\frac{H^{\mathrm {ig} }-H}{RT} =\int_{V}^{\infty }\left[T\left({\frac{\partial Z}{\partial T}}\right)_{V}\right]{\frac{dV}{V}} + 1-Z
		\end{equation}
	}

	The Peng–Robinson equation of state relates the three interdependent state properties pressure $P$, temperature $T$, and molar volume $V_m$. From the state properties ($P$, $V_m$, $T$), one may compute the departure function for enthalpy per mole (denoted $h$) as presented by \citet{Gmehling2019} or \citet{Elliott2011}:

	{\footnotesize
		\begin{equation}
			h_{T,P}-h_{T,P}^{\mathrm {ideal} }=RT_{C}\left[T_{r}(Z-1)-2.078(1+\kappa ){\sqrt {\alpha }}\ln \left({\frac {Z+2.414B}{Z-0.414B}}\right)\right]
		\end{equation}
	}
